\section{Context Fluency: An Emerging Skill}

Our case study and the broader practitioner literature suggest that effective
agentic coding requires a skill set distinct from both traditional
programming and prompt engineering. We propose the term \emph{context
fluency} to describe this emerging capability: the ability to create rich,
structured context that AI agents can act on effectively. Context fluency
encompasses capturing domain expertise, value judgments, and design intent in
machine-legible form---not as one-shot prompts, but as persistent
informational environments within which agents operate. This distinction is
critical. Prompt engineering optimizes individual instructions to a model;
context fluency is \emph{upstream} of prompting, concerned with the entire
informational architecture that surrounds agent
execution~\cite{anthropic2025context, gartner2025context}.

We identify four components of context fluency as practiced in our case study
and observed in adjacent practitioner work:
\begin{description}
  \item[Decomposition:] Breaking problems into discrete, parallelizable
    tasks with clear boundaries---enabling multiple agents to work
    concurrently without coordination overhead.
  \item[Specification:] Describing not just \emph{what} to build but
    \emph{why}, so that agents can make aligned micro-decisions at
    implementation time without requiring human intervention.
  \item[Constraint definition:] Knowing what to exclude, simplify, or
    defer---scope management as a first-class concern. In our hackathon case,
    constraining the product to a single assignment type was as important as
    any feature we included.
  \item[Domain encoding:] Capturing lived experience, value judgments, and
    tacit knowledge~\cite{polanyi1966tacit} that agents cannot generate
    independently. The twenty minutes spent dictating pedagogical intuitions
    about productive struggle produced the context that made the Socratic
    tutor's behavior correct on first implementation.
\end{description}

These components map to established theoretical frameworks. Backward
design~\cite{wiggins1998understanding} contributes the principle of
outcome-driven preparation: defining what success looks like before
designing the path to reach it. Polanyi's work on tacit
knowledge~\cite{polanyi1958personal} frames the externalization challenge:
practitioners ``know more than they can tell,'' and context fluency is
precisely the discipline of telling it anyway, in a form that agents can
consume. We also observe a strong connection to pedagogical scaffolding.
The skill set of context fluency closely resembles lesson planning in
education---creating conditions where the learner (or agent) can do
meaningful work independently, because the environment has been structured
in advance. This connection between teaching and context engineering is not
coincidental: both are fundamentally about enabling others to act effectively
within a defined informational context.

If context fluency is a distinct and identifiable skill, the implications
extend beyond individual practice. For developer education, it suggests that
curricula should cultivate specification, decomposition, and domain-encoding
abilities alongside traditional programming competencies. For hiring, it
implies that practitioners with strong domain knowledge and pedagogical
instincts---the ability to structure information for consumption by
others---may be disproportionately effective in agentic
workflows~\cite{dakhel2025humanai}. For tool design, it reframes preparation
aids such as structured task systems~\cite{yegge2025beads,
emanuel2025beadsrust}, specification frameworks~\cite{github2025speckit},
and context engineering platforms~\cite{anthropic2025context} as essential
infrastructure rather than optional enhancements.
