\section{Introduction}

AI coding agents have become remarkably capable at code generation.
Controlled studies of GitHub Copilot report productivity gains of 21--55\%
across programming tasks~\cite{peng2023copilot}, and the broader ecosystem
of autonomous agents can now scaffold entire applications from natural-language
descriptions. Yet the dominant workflow pattern for leveraging these
capabilities---what Karpathy~\cite{karpathy2025vibe} termed ``vibe coding''---prioritizes
speed of implementation over deliberate preparation. The developer describes
an intent, the agent produces code, and misalignments are resolved through
iterative correction. This creates a systematic alignment problem: agents
operating without sufficient context produce code that requires extensive
rework. Veracode reports that 45\% of AI-generated code contains security
flaws~\cite{veracode2025ai}, and practitioners consistently observe that
debugging misaligned agent output can consume substantial
development time~\cite{huntley2025ralph, yegge2025beads}. The bottleneck in
agentic coding is not code generation but alignment between developer intent
and agent output---that is, the degree to which agent-produced artifacts
exhibit specification compliance (planned features built as specified),
architectural fidelity (design decisions honored across the codebase),
and low rework ratio (corrective commits as a proportion of total commits).

We argue that this alignment problem is fundamentally a \emph{preparation}
problem. Drawing on the culinary concept of \emph{mise en place}---a French
term meaning ``everything in its place''---we propose a preparation-first
methodology for agentic coding. In professional kitchens, thorough
preparation before service enables fluid execution: every ingredient is
measured, every tool is positioned, so that once cooking begins, the chef's
hands never pause to search. The same principle applies to agent
orchestration. When domain expertise, design intent, and task boundaries are
externalized into structured artifacts \emph{before} agents begin writing
code, the resulting implementation is more aligned, more coherent, and less
costly to verify. Our methodology formalizes this preparation phase into
three distinct steps: contextual grounding, where tacit knowledge and domain
expertise are captured in machine-legible documents; collaborative
specification, where human-agent dialogue produces detailed design artifacts;
and task decomposition, where specifications are converted into structured,
dependency-aware work records.

This paper makes four contributions:
\begin{enumerate}
  \item A three-phase mise en place methodology for agentic coding, grounded
    in backward design~\cite{wiggins1998understanding} and tacit knowledge
    externalization~\cite{polanyi1966tacit}.
  \item A case study applying this methodology in a competitive hackathon,
    with quantitative data on preparation artifacts and implementation
    outcomes.
  \item The concept of \emph{context fluency} as an emerging developer
    skill---the ability to create rich, structured context that AI agents can
    act on.
  \item A research agenda with five open questions for empirical validation
    of preparation-phase methodologies.
\end{enumerate}
