\section{Background and Related Work}
\label{sec:related}

Early empirical studies of AI-assisted coding reported substantial productivity
gains: Peng et al.\ found that developers using GitHub Copilot completed tasks
21--55\% faster~\cite{peng2023copilot}. However, speed has not translated cleanly
into quality. Veracode's 2025 analysis found that 45\% of AI-generated code
contains security flaws~\cite{veracode2025ai}, while Mozannar et al.\ showed
that developers' mental models frequently diverge from actual AI behavior, leading
to inappropriate acceptance of incorrect suggestions~\cite{developer2023mental}.
Vasconcelos et al.\ demonstrated that trust in AI code generation depends on
context-aware explanations that current tools rarely
provide~\cite{facct2024trust}. Together, these findings point to a productivity paradox:
AI agents accelerate output while simultaneously increasing the surface area for
rework and defects. The bottleneck in agentic coding, we argue, is not code
generation but \emph{alignment}---ensuring that what agents build matches what
practitioners intend.

\subsection{Prompt and Context Engineering}

A substantial body of work studies how to elicit better behavior from
large language models through careful instruction. Foundational results
established that LLMs perform new tasks from in-context
examples~\cite{brown2020gpt3}, and that intermediate reasoning steps
improve task accuracy~\cite{wei2022chain}. Liu et al.'s
survey~\cite{liu2023prompt} catalogues a maturing taxonomy of prompting
techniques. More recently, the discourse has shifted from prompt
engineering to \emph{context engineering}: designing the complete
informational environment surrounding an LLM~\cite{anthropic2025context,
gartner2025context}. Alomar et al.\ extend this thread with
\emph{promptware engineering}~\cite{alomar2025promptware}, and Dakhel
et al.'s taxonomy identifies eleven distinct human-AI interaction
types~\cite{dakhel2025humanai}. MEP operates at a different scope from
prompt engineering: prompt engineering tunes individual model
invocations, whereas MEP structures the workflow-level artifacts
(specifications, decompositions, externalized tacit knowledge) that
precede and constrain those invocations. Both shape what the model
receives, but at different granularities.

\subsection{Preparation Methodologies}

Several preparation-oriented methodologies anticipate aspects of our
proposal. Horthy's Research-Plan-Implement (RPI)
methodology~\cite{horthy2025rpi} most directly approaches this work,
requiring practitioners to complete discrete research and planning
phases before any code is generated. MEP differs from RPI by
formalizing tacit knowledge externalization in Phase~1 and insisting on
backward-design sequencing where all three phases complete before
implementation begins. GitHub's Spec Kit~\cite{github2025speckit}
formalizes specification-driven development, where a structured
specification artifact governs agent behavior throughout
implementation. These contributions recognize that what happens
\emph{before} agent execution shapes outcomes, but none offers a
unified, theoretically grounded preparation methodology that
integrates tacit knowledge externalization, collaborative
specification, and dependency-aware decomposition into one phase-gated
sequence.

\subsection{Practitioner Tooling and Theoretical Grounding}

A parallel stream of practitioner innovation addresses preparation
through tooling. Karpathy coined ``vibe coding''~\cite{karpathy2025vibe}
for a minimal-preparation workflow scoped to throwaway projects, yet the
term came to characterize the dominant mode of AI-assisted development.
At the opposite end, Huntley's Ralph Loop~\cite{huntley2025ralph}
provides fine-grained iterative task delegation, and Yegge's Beads
framework~\cite{yegge2025beads} addresses agent memory through
structured task records, extended by Emanuel's Rust
reimplementation~\cite{emanuel2025beadsrust}. These tools solve specific
problems---iteration, memory, orchestration---but assume that
appropriate context already exists. MEP draws on two theoretical
frameworks. Wiggins and McTighe's
\emph{backward design}~\cite{wiggins1998understanding} holds that
effective instruction begins with desired outcomes, then works backward
through assessment criteria---we apply this principle by defining
agent outputs and success criteria before designing context. Polanyi's
tacit knowledge---``we can know more than we can
tell''~\cite{polanyi1966tacit, polanyi1958personal}---frames preparation
as a knowledge externalization problem; Nonaka and Takeuchi's SECI
model~\cite{nonaka1995knowledge} formalizes this tacit-to-explicit
conversion as a repeatable organizational process, grounding Phase~1.

\subsection{Positioning: What is New}
\label{sec:novelty}

The ingredients of MEP are individually well-established; our
contribution is integrative and scope-specific. Against
\emph{spec-driven development}~\cite{github2025speckit}, MEP adds
Phase~1 for externalizing tacit, value-laden knowledge that
specifications alone do not capture. Against \emph{prompt
engineering}~\cite{brown2020gpt3, wei2022chain, liu2023prompt}, MEP
operates at workflow rather than invocation scope. Against
\emph{iterative or vibe-coding workflows}~\cite{karpathy2025vibe,
huntley2025ralph}, MEP front-loads alignment work that iterative flows
pay incrementally as rework. Agentic workflows need this preparation
because agents lack the tacit context human collaborators carry and
cannot iterate cheaply without expensive re-generation.
