\section{Limitations, Research Agenda, and Conclusion}
\label{sec:agenda}

\subsection{Limitations}
\label{sec:limitations}

The empirical claims in this paper rest on a single hackathon, a single
practitioner, and no control group. We acknowledge these limits
directly. (i)~\textbf{Single hackathon, single case study.} Findings
are illustrative, not generalizable; the five-hour competitive setting
is not representative of multi-month software development, and a
different practitioner, project, or AI model may produce different
results. (ii)~\textbf{No control group, no iteration-first comparison.}
We did not run a paired condition in which the same practitioner
attempted the same project without preparation, nor did we instrument
the other 11 teams. The 5.7:1 preparation-to-execution ratio and
near-zero architectural rework are reported observations, not effects
estimated against a baseline. (iii)~\textbf{Operator expertise as a
confound.} The practitioner's dual professional background in education
and software engineering likely contributed to the quality of
contextual grounding---particularly the pedagogical encoding that
informed the Socratic tutor---so we cannot separate the methodology's
contribution from operator expertise. The qualitative team-variation
sketch in Section~\ref{sec:teamvariation} adds context but does not
constitute controlled comparison. We present these results as
suggestive evidence that MEP merits further investigation, and
explicitly frame the empirical questions as future work below.

\subsection{Research Agenda}

MEP as described here rests on a single case study and conceptual
argument. Empirical validation across practitioners, domains, and agent
models is essential. We propose five research questions and three
concrete research lines that follow directly from the limitations
above:

\begin{description}
  \item[RQ1:] \emph{Does deliberate preparation reduce agent misalignment
    compared to iterative development, and at what cost-benefit threshold?}
    A \textbf{comparative study} pairing preparation-first and
    iteration-first workflows (preparation-first vs.\ iteration-first)
    on matched tasks would establish whether the upfront investment
    yields net savings; this directly addresses Limitation~(ii).

  \item[RQ2:] \emph{What constitutes ``sufficient'' context for agentic
    work---can we define stopping criteria for the preparation phase?}
    A saturation-style analysis measuring alignment quality as a function of
    context volume would help practitioners know when to stop preparing.

  \item[RQ3:] \emph{How does context fluency vary across practitioners,
    domains, and agent models?} A
    \textbf{multi-team replication study} across diverse codebases and
    practitioners---generalizing the team-variation sketch in
    Section~\ref{sec:teamvariation}---could reveal whether context
    fluency is trainable or a function of prior domain
    knowledge~\cite{polanyi1958personal}, addressing
    Limitation~(iii).

  \item[RQ4:] \emph{Can MEP scale beyond prototyping to sustained,
    multi-month development?} A \textbf{longitudinal study} of
    context-fluency development on multi-month projects would test the
    methodology's applicability at scale~\cite{dakhel2025humanai} and
    address Limitation~(i).

  \item[RQ5:] \emph{What is the relationship between preparation time and
    implementation quality across project complexities?} A multi-project
    dataset correlating preparation artifacts with implementation metrics
    could yield quantitative models of the
    preparation-quality relationship~\cite{alomar2025promptware}.
\end{description}

The three named research lines---comparative, replication, and
longitudinal---convert the empirical concerns raised by the
single-case design into a falsifiable program. RQ1 is the most
immediately tractable, addressable with paired tasks and a modest
participant pool.

\subsection{Artifact Availability}
\label{sec:artifact}

To support the SIGSOFT Open Science Policy and enable replication, we
release a research artifact bundle containing the anonymized hackathon
data (timing, commit, and bead records reported in this paper), the
prompts and agent configurations used, and the context-engineering
scaffolds (CLAUDE.md, beads schema, planning templates) that
constitute the MEP harness. The bundle is licensed CC-BY~4.0 (data) and
MIT (code). It is archived on Zenodo at \url{https://zenodo.org/[TBD]}
(DOI to be inserted in the camera-ready upload).

\subsection{Conclusion}

MEP, as presented here, is a starting framework rather
than a finished methodology---a formalization of practices that
experienced practitioners already converge on
independently~\cite{huntley2025ralph, horthy2025rpi,
yegge2025beads}. The methodology formalizes what Nonaka and
Takeuchi~\cite{nonaka1995knowledge} term externalization---converting
tacit practitioner knowledge into explicit artifacts that agents can
consume. Our case study suggests that deliberate preparation
can compress implementation time and reduce architectural
misalignment, but we claim no generalizability from a single
hackathon. What we do claim is that the alignment problem in agentic
coding is fundamentally a preparation problem, and that the field
would benefit from treating the work before agents write code as a
first-class engineering discipline.

Context fluency offers a lens for understanding emerging practitioner skills---skills
that bridge domain expertise, pedagogical design, and specification
craft~\cite{anthropic2025context, github2025speckit}. We invite empirical
validation, replication across domains, and the development of tools that support
preparation-phase work as the field moves from ad-hoc agent interaction toward
disciplined context engineering.
